\mytitle

### 2020-07-17 paper.tex
Edits through the rest of the document.

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site and information about the wind patterns, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference between turbines. Given the pairwise wake interactions, it is natural
%While a number of heuristic and exact techniques have been proposed to solve WFLO, it is common
to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization approaches. However, state-of-the-art solutions often trade-off between efficiency (providing quick designs) and accuracy (guaranteeing optimal energy solutions). In this work, we aim to find a balancing approach to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when accuracy is the goal. To this end, we exploit recent advances in both optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
We 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and specialized QUBO hardware
%demonstrates the strengths of the two models. 
%Our results
show that our quadratic framework achieves fast, high-quality solutions 
that
improve the state of the art and strike a balance between efficiency and optimality. In particular, the QUBO model delivers high quality solutions in a few seconds while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

The remainder of this paper is structured as follows. Section~\ref{sec:related} presents our wind farm model and a literature review of exact and approximate
optimization solutions to the layout problem. The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO} and Section~\ref{sec:eval} demonstrates the value of our approach via a thorough empirical evaluation.
%of the proposed methods.
In Section~\ref{sec:discussion}, we discuss the results and their practical implications.
Lastly, Section~\ref{sec:conclusion} provides concluding remarks and directions for future work.

@@

In this section, we 
present an empirical evaluation of the two quadratic models based on 
12 standard WFLO instances.
Our main results are: \begin{itemize}
	\item For given time limit, the Digital Annealer (DA) achieves a solution quality as good or as better than existing state-of-the-art exact methods on 10 of 12 instances.
		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.  
	\item Given sufficient runtime, solving the \qcls{} model with commercial software yields strong results on most of the instances. Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard ones. %  \todo{Arik: Is this correct? Table 2 shows DA@10seconds with the best solution found for 7/12 instances and \qcls{}grb@3600 for 6/12. And 4/6 vs. 3/6 for the WR36). Are you comparing \qcls{}grb with \qcssgrb and ILP-LS? We seem to be underplaying DA performance in this comment. Chris: I changed to competitive results. You're right, I'd say that the DA is not `losing' to \qcls{}grb.}  
\end{itemize}

@@

Our results show empirically that the quadratic programming models that we
introduced in Section~\ref{sec:QUBO4WFLO} (\qcls{} and \quls{}) are useful in two ways.
First, the two methods are powerful approaches to solving WFLO. Specifically, \qcls{} yields the best performance (after 3600 seconds) for most problem instances
and especially,
for the hard instances with 36 wind directions. In addition, the runtime performance of \qcls{} is second
only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the \qulsda{} approach. \qulsda{} yields the best results in 7 out of 12 
instances, shows competitive performance on the other instances, and outperforms all methods in terms of performance in short runtimes (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second).

@@

In this work, we presented a novel quadratic programming framework
for solving 
wind farm layout optimization. 
The framework is embodied in two models: constrained (\qcls) and unconstrained binary
 (\quls) quadratic programs. The former can be used to solve WFLO using advanced optimization software,
while the latter can be mounted onto specialized hardware that can quickly (approximately) solve WFLO. We used the constrained quadratic formulation to prove 
that the computational complexity of WFLO is $\mathcal{NP}$-hard. Nonetheless,
our empirical evaluation shows that when exact solvers like Gurobi~\cite{gurobi} are
coupled with our \qcls{} and \quls{} models they are able to solve small WFLO instances to optimality in less than 3600 seconds.
We also show that when an advanced hardware is used with the QUBO representation, large WFLO instances can be solved quickly with no significant degradation in the total energy compared to exact methods.
This runtime improvement and state-of-the-art performance in terms of the total energy of the resulting solutions makes our quadratic framework useful for designing new wind farms, while varying multiple input parameters.

>>>>

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site and information about the wind patterns, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference between turbines. Given the pairwise wake interactions, it is natural
%While a number of heuristic and exact techniques have been proposed to solve WFLO, it is common
to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization approaches. However, state-of-the-art solutions often trade-off between efficiency (providing quick designs) and accuracy (guaranteeing optimal energy solutions). In this work, we aim to find a balancing approach to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when accuracy is the goal. To this end, we exploit recent advances in both optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
We 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough numerical 
evaluation using a commercial solver and specialized QUBO hardware
%demonstrates the strengths of the two models. 
%Our results
show that our quadratic framework achieves fast, high-quality solutions 
that
improve the state of the art and strike a balance between efficiency and optimality. In particular, the QUBO model delivers high quality solutions in a few seconds while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

The remainder of this paper is structured as follows. Section~\ref{sec:related} presents our wind farm model and a literature review of exact and approximate
optimization solutions to the layout problem. The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO} and Section~\ref{sec:eval} demonstrates the value of our approach via a thorough  numerical evaluation.
%of the proposed methods.
In Section~\ref{sec:discussion}, we discuss the results and their practical implications.
Lastly, Section~\ref{sec:conclusion} provides concluding remarks and directions for future work.

@@

In this section, we 
present an numerical evaluation of the two quadratic models based on 
12 standard WFLO instances.
Our main results are: \begin{itemize}
	\item For given time limit, the Digital Annealer (DA) achieves a solution quality as good or as better than existing state-of-the-art exact methods on 10 of 12 instances.
		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.  
	\item Given sufficient runtime, solving the \qcls{} model with commercial software yields strong results on most of the instances. Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard ones. %  \todo{Arik: Is this correct? Table 2 shows DA@10seconds with the best solution found for 7/12 instances and \qcls{}grb@3600 for 6/12. And 4/6 vs. 3/6 for the WR36). Are you comparing \qcls{}grb with \qcssgrb and ILP-LS? We seem to be underplaying DA performance in this comment. Chris: I changed to competitive results. You're right, I'd say that the DA is not `losing' to \qcls{}grb.}  
\end{itemize}

@@

Our results show numerically that the quadratic programming models that we
introduced in Section~\ref{sec:QUBO4WFLO} (\qcls{} and \quls{}) are useful in two ways.
First, the two methods are powerful approaches to solving WFLO. Specifically, \qcls{} yields the best performance (after 3600 seconds) for most problem instances
and especially,
for the hard instances with 36 wind directions. In addition, the runtime performance of \qcls{} is second
only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the \qulsda{} approach. \qulsda{} yields the best results in 7 out of 12 
instances, shows competitive performance on the other instances, and outperforms all methods in terms of performance in short runtimes (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second).

@@

In this work, we presented a novel quadratic programming framework
for solving 
wind farm layout optimization. 
The framework is embodied in two models: constrained (\qcls) and unconstrained binary
 (\quls) quadratic programs. The former can be used to solve WFLO using advanced optimization software,
while the latter can be mounted onto specialized hardware that can quickly (approximately) solve WFLO. We used the constrained quadratic formulation to prove 
that the computational complexity of WFLO is $\mathcal{NP}$-hard. Nonetheless,
our numerical evaluation shows that when exact solvers like Gurobi~\cite{gurobi} are
coupled with our \qcls{} and \quls{} models they are able to solve small WFLO instances to optimality in less than 3600 seconds.
We also show that when an advanced hardware is used with the QUBO representation, large WFLO instances can be solved quickly with no significant degradation in the total energy compared to exact methods.
This runtime improvement and state-of-the-art performance in terms of the total energy of the resulting solutions makes our quadratic framework useful for designing new wind farms, while varying multiple input parameters.
